PAMELA: an annotation-based Java Modeling Framework

### 2021-02-08 6-Experimentations.tex
Improving related work.

The main results of the project are a requirements modeling multi-views language, its associated design process and the development of an opensource platform called Formod\cite{FormodWebSite}, built using Openflexo infrastructure and PAMELA framework. The requirements modeling language is based on KAOS for goal modeling and SysML for the structural part of a system. The associated domain modeling language, used to describe system domain knowledge, extends the two ontology languages OWL and PLIB. The graphical notations are then translated into Event-B, a formal specification method supported by verification tools.

>>>>

The main results of the project are a requirements modeling multi-views language, its associated design process and the development of an opensource platform called Formod\cite{FormodWebSite}, built using Openflexo infrastructure and PAMELA framework. \todo{Say how and where PAMELA was used} The requirements modeling language is based on KAOS for goal modeling and SysML for the structural part of a system. The associated domain modeling language, used to describe system domain knowledge, extends the two ontology languages OWL and PLIB. The graphical notations are then translated into Event-B, a formal specification method supported by verification tools.

\todo{Remark that it is industry ready}

### 2021-02-08 related.tex
Improving related work.

When the observed artifact is code (e.g., for runtime analysis or verification of properties) we could see our approach as similar to some contributions appertaining to the models@runtime topic~\cite{bencomo2019models}. Indeed models@runtime approaches often rely  on the use of the the reflection architectural pattern~\cite{buschmann2008pattern} in order to separate core application logic from a metalevel that contains information about properties. This is also the case of PAMELA. However the focus of most models@runtime approaches is not on easing the development process but on enabling the analysis and/or the adaptation/evolution of ever running applications. Examples of such approaches are AC-Contrat~\cite{accontract} focusing in the Run-time verification of properties for context-aware applications. Ramses~\cite{ramses} focusing on dynamic adaptations,~\cite{denker2010modeling} focusing on feature analysis and~\cite{song2010applying}, which wraps running system in standard UML-like models in order to perform analysis and management task by using off-the-shelf MDE tools and techniques.

More similar to PAMELA, FAME \cite{kuhn2008fame} is  a library that keeps metamodels
accessible and adaptable at runtime. Basically, FAME attaches meta-information
to the objects of a running application permitting a uniform manipulation of
both models and system objects at runtime.

More recently, in \cite{boronat2019} the authors construct and
maintain at runtime model-based views on the data manipulated in
object-oriented code. Different to the aforementioned approaches, PAMELA does
not focus in reflecting and manipulating running systems but in: 1) extending
their behaviors; and 2) maintaining a continuous cycle of metamodeling so that
models and corresponding code remain synchronized without the need for code
generation processes. In a different approach, UMPLE
\cite{lethbridge2016merging} mixes programming and modeling by integrating UML
constructs into languages such as Java. However, they use code generation for
the runtime part of the system.

>>>>

When the observed artifact is code (e.g., for runtime analysis or verification of properties) we could see our approach as similar to some contributions appertaining to the models@runtime topic~\cite{bencomo2019models}. Indeed models@runtime approaches often rely  on the use of the the reflection architectural pattern~\cite{buschmann2008pattern} in order to separate core application logic from a metalevel that contains information about properties. This is also the case of PAMELA. 
%However the focus of most models@runtime approaches is not on easing the development process but on enabling the analysis and/or the adaptation/evolution of ever running applications. 
Examples of such approaches are AC-Contrat~\cite{accontract} focusing in the Run-time verification of properties for context-aware applications. Ramses~\cite{ramses} focusing on dynamic adaptations,~\cite{denker2010modeling} focusing on feature analysis and~\cite{song2010applying}, which wraps running system in standard UML-like models in order to perform analysis and management task by using off-the-shelf MDE tools and techniques. More similar to PAMELA, FAME \cite{kuhn2008fame} is a polyglot library that keeps metamodels accessible and adaptable at runtime. Basically, FAME attaches meta-information to the objects of a running application permitting a uniform manipulation of both models and system objects at runtime. Synchronization of changes between models and code are however limited for languages such as Java. In~\cite{cavarle2016dynamic} the authors present an approach to keep the (bi-directional) synchronization between feature-based models and generated code alternatives. They rely on Pharo\cite{pharo} reflective capabilities. 
More recently, in \cite{boronat2019} the authors construct and maintain at runtime model-based views on the data manipulated in object-oriented code.

Different to the aforementioned approaches, the focus of PAMELA is not on adapting or observing ever running programs (although it can be used for the verification of runtime properties as demonstrated in~\cite{silva20}) but on: 1) providing a mechanism to blend coding and metamodeling so that the code and the metamodel may be built incrementally without the need for code generation and thus, avoiding round-tripping issues; and 2) providing default implementations for frequently used concepts in order to ease development. In that sense PAMELA can be seen as a mix between classical CASE tools and the aforementioned models@runtime approaches.

In a different approach, UMPLE\cite{lethbridge2016merging} mixes programming and modeling by integrating UML
constructs into languages such as Java. However, they use code generation for
the runtime part of the system.
